An event-sourced and distributed database with IPLD and materialized views

### 2019-09-20 draft.tex
refs: adds more refs + modifies gitignore
Signed-off-by: Carson Farmer <carson.farmer@gmail.com>

\subsubsection{Logical Clocks}

@@

Cryptographically linked events can also represent a clock. One such example is called the Merkle-Clock \cite{sanjuanMerkleCRDTs2019}. The Merkle-Clock relies on properties of a Merkle-DAG to provide strict partial ordering between events, which has limitations \cite[sec 4.3]{sanjuanMerkleCRDTs2019}:

@@

As discussed above, IPFS uses the cryptographic hash of a given piece of content to define its content address (see [ref] for details on this process). However, in order to provide standards for accessing content-addressable data (on the web or elsewhere), it is necessary to define a common format or specification. In IPFS (and other projects [ref]), this common data format is called Interplanetary Linked Data (IPLD) [ref]. As the name suggests, IPLD is based on principals of linked data [ref, ref] with the added capabilities of a content-addressing storage network.

IPLD is used to publish linked data (subject-predicate-object triples in the linked data world [ref]) spread across different hosts, and where everything (entities, predicates, data sources) uses content addresses as unique identifiers [ref]. To form its structure, IPLD implements a Merkle DAG, or directed acyclic graph. This allows all hash-linked data structures to be treated using a unified data model, analogous to linked data in the Semantic Web sense [ref]. In practice, IPLD is represented as objects, each with `Data` and `Links` fields, where `Data` can be a small blob of unstructured, arbitrary binary data, and `Links` is an array of links to other IPLD objects.

@@

A Merkle-Clock is a Merkle-DAG that represents a sequence of events. In other words, a Merkle-Clock is an append-only log [ref and ipfs-log]. When implemented on IPFS (or an equivalent network where content can be cryptographically addressed and fetched), Merkle-Clocks provide a number of benefits for data synchronization between replicas [ref merkle paper section 4.3]:

@@

On the downside, and as seen above (section 2.2.4.), a Merkle-Clock cannot order divergent heads (or roots):

@@

In order to reduce the likelihood of divergent heads, all replicas should be perfectly connected and be able to fetch all events and linkages in the Merkle-Clock. On real networks with many often offline replicas (mobile and IoT devices, laptops, etc.), these conditions are rarely met.

Based on these observations, using a single Merkle-Clock to synchronize replicas can be problematic.

@@

The libp2p project provides a robust protocol communication stack. IPFS and a growing list of other projects [ref] are building on top of libp2p. Libp2p solves a number of challenges that are distinct to peer-to-peer (p2p) networks. A comprehensive coverage of networking issues in P2P systems is out of the scope for this paper, however, some core challenges that libp2p helps to address include NAT traversal, peer discovery and handshake protocols, and even encryption and transport security — libp2p supports both unencrypted (e.g. TCP, UDP) and encrypted protocols (e.g. TLS, Noise) — among others. Libp2p uses the concept of a multiaddress to address peers on a network, which essentially models network addresses as arbitrary encapsulations of protocols [ref, ref]. In addition to “transport layer” modules, libp2p provides several tools for sharing and/or disseminating data over a p2p network.

>>>>

\subsubsection{Logical Clocks}\label{sec:LogicalClocks}

@@

Cryptographically linked events can also represent a clock. One such example is called the Merkle-Clock \cite{sanjuanMerkleCRDTs2019}. The Merkle-Clock relies on properties of a Merkle-DAG to provide strict partial ordering between events, which has limitations \cite[sec. 4.3]{sanjuanMerkleCRDTs2019}:

@@

As discussed above, IPFS uses the cryptographic hash of a given piece of content to define its content address (see \cite{benetIpfscontentAddressedVersioned2014} for details on this process). However, in order to provide standards for accessing content-addressable data (on the web or elsewhere), it is necessary to define a common format or specification. In IPFS (and others \cite[e.g.,]{protocollabsFilecoinDecentralizedStorage2017} MORE NEEDED), this common data format is called Interplanetary Linked Data (IPLD)\footnote{https://ipld.io/}. As the name suggests, IPLD is based on principals of linked data \cite{timberners-leeLinkedData2009,bizerLinkedDataStory2011} with the added capabilities of a content-addressing storage network.

IPLD is used to publish linked data (subject-predicate-object triples in the linked data world \cite{heathLinkedDataEvolving2011}) spread across different hosts, and where everything (entities, predicates, data sources) uses content addresses as unique identifiers. To form its structure, IPLD implements a Merkle DAG, or directed acyclic graph\footnote{For reference, the Bitcoin Blockchain is a DAG, so is a Git version history.}. This allows all hash-linked data structures to be treated using a unified data model, analogous to linked data in the Semantic Web sense \cite{brendanobrienDeterministicQueryingDistributed2017}. In practice, IPLD is represented as objects, each with \texttt{Data} and \texttt{Links} fields, where \texttt{Data} can be a small blob of unstructured, arbitrary binary data, and \texttt{Links} is an array of links to other IPLD objects.

@@

A Merkle-Clock is a Merkle-DAG that represents a sequence of events. In other words, a Merkle-Clock is an append-only log \cite{sanjuanMerkleCRDTs2019}. When implemented on IPFS (or an equivalent network where content can be cryptographically addressed and fetched), Merkle-Clocks provide a number of benefits for data synchronization between replicas \cite[sec. 4.3]{sanjuanMerkleCRDTs2019}:

@@

On the downside, and as seen above (section \ref{sec:LogicalClocks}), a Merkle-Clock cannot order divergent heads (or roots):

@@

In order to reduce the likelihood of divergent heads, all replicas should be perfectly connected and be able to fetch all events and linkages in the Merkle-Clock. On real networks with many often offline replicas (mobile and IoT devices, laptops, etc.), these conditions are rarely met. Based on these observations, using a single Merkle-Clock to synchronize replicas can be problematic.

@@

The libp2p project provides a robust protocol communication stack. IPFS and a growing list of other projects [MISSING] are building on top of libp2p. Libp2p solves a number of challenges that are distinct to peer-to-peer (p2p) networks. A comprehensive coverage of networking issues in P2P systems is out of the scope for this paper, however, some core challenges that libp2p helps to address include NAT traversal, peer discovery and handshake protocols, and even encryption and transport security — libp2p supports both unencrypted (e.g. TCP, UDP) and encrypted protocols (e.g. TLS, Noise) — among others. Libp2p uses the concept of a multiaddress to address peers on a network, which essentially models network addresses as arbitrary encapsulations of protocols [ref, ref]. In addition to “transport layer” modules, libp2p provides several tools for sharing and/or disseminating data over a p2p network.
